---
title: "Data analysis via single and multilevel regression"
author: "Oscar Garcia"
date: "2024-09-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary
The following R.Markdown script (work in progress) aims to conduct a series of single and multilevel regression analyses. It is part of a larger study in progress exploring national identity typologies.

As independent variables, it focuses on:
- Core demographic variables (level-1, individual): age (agerange), education (edulvl) and quintile income (quintinc).
- Immigration rate (level-2, country) (immigrate).

As dependent variables, it leverages:
- Difference between ethnic and cultural national identity scores (eth_vs_cult).
- Overall national identity score (nationalidentity).

On a third step, it tests the effect of immigrate, eth_vs_cult and nationalidentity on rejection of multiculturalism.

#1. Preparatory tasks

##1.a) Reset space and charge libraries

```{r, echo=FALSE}
rm(list=ls()) # Clean workspace
options(scipen=10000) # Set scientific notation
library(tidyverse)
library(gridExtra)
library(stargazer) # For regression tables
library(lme4) # For mixed effects models
library(MuMIn) # For Conditional R² in mixed effects models
library(sjPlot) # For creating mixed effects' regression tables
library(sjmisc)
library(sjlabelled)

```

##1.b) Open dataset 

```{r}
df_nationalism <- read.csv("C:/Users/oscar/Documents/Projects/Paper- Lv2EffectsOnNationalIdentity/Quantitative Analysis/Data/df_nationalism.csv", header = TRUE, sep = ",", encoding = "UTF-8")

```

##1.c) Prepare the data

###1.c) 1.- Delete observations with at least one parent born outside of country or, where this is not available, the respondant born outside of country

```{r}
df_nationalism <- df_nationalism %>%
  filter(countryname %in% c("Mexico", "Australia", "Japan", "Netherlands", "Norway", "South Korea", "Uruguay") & immigrant == 1 | (!(countryname %in% c("Mexico", "Australia", "Japan", "Netherlands", "Norway", "South Korea", "Uruguay") & parentforeign == 1)))

```


###1.c) 2.- Delete observations with missing values across any of the key variables

```{r}
# Delete missing cases 
df_nationalism <- df_nationalism %>%
  filter(complete.cases(eth_vs_cult, 
                        edulvl, 
                        quintinc,
                        l2immigrate, 
                        ougrcusttrad,
                        nationalidentity,
                        agerange,
                        gender,
                        age
                        ))

```

#2. Descriptive analyses
##2. a) Create descriptive statistics table of groups

```{r}
# Summarize data by group
summary_by_group <- df_nationalism %>%
  # Filter by groups defined in the vector
  filter(countryname %in% df_nationalism$countryname) %>%
  # Group by country and year
  group_by(countryname) %>%
  # Summarize with desired statistics, handling NAs with na.rm=TRUE
  summarise(
    N = n(),  # Count observations (number of individuals)
    Mean_Age = round(mean(age, na.rm = TRUE), digits = 2),  # Mean age (ignoring NAs)
    SD_Age = round(sd(age, na.rm = TRUE), digits = 2),  # Standard deviation of age (ignoring NAs)
    Pct_Female = round(mean(gender == 1, na.rm = TRUE), digits = 4) * 100  # Percent female (ignoring NAs)
  )

# Summarize data by group
summary_total <- df_nationalism %>%
  # Summarize with desired statistics, handling NAs with na.rm=TRUE
  summarise(
    N = n(),  # Count observations (number of individuals)
    Mean_Age = round(mean(age, na.rm = TRUE), digits = 2),  # Mean age (ignoring NAs)
    SD_Age = round(sd(age, na.rm = TRUE), digits = 2),  # Standard deviation of age (ignoring NAs)
    Pct_Female = round(mean(gender == 1, na.rm = TRUE), digits = 4) * 100  # Percent female (ignoring NAs)
  )

# Add a "countryname" column to summary_total
summary_total <- summary_total %>%
  mutate(countryname = "Total")

# Combine the two tables
combined_summary <- bind_rows(summary_by_group, summary_total)

# Print the combined summary table
print(combined_summary)

```

##2. b) Create subset for analysis and perform some descriptive analyses to analyse outcome variables
###2. b) 1.- Difference between ethnic and cultural

```{r}
summary(df_nationalism$eth_vs_cult)

sd(df_nationalism$eth_vs_cult)

hist(df_nationalism$eth_vs_cult, breaks = 20, main = "Histogram of Ethnic vs Cultural Nationalism", xlab = "Ethnic - Cultural Nationalism", col = "lightblue")

# Create a weighted, stacked histogram with ggplot2
ggplot(df_nationalism, aes(x = eth_vs_cult, fill = countryname, weight = polweight)) +
  geom_histogram(bins = 20, color = "black") +
  scale_fill_manual(values = rainbow(length(unique(df_nationalism$countryname)))) +
  labs(title = "Weighted Histogram of Ethnic vs Cultural Nationalism",
       x = "Ethnic - Cultural Nationalism",
       y = "Weighted Frequency") +
  theme_minimal()


summary(df_nationalism$polweight)
plot(density(df_nationalism$eth_vs_cult), main = "Density Plot of Ethnic vs Cultural Nationalism", xlab = "Ethnic - Cultural Nationalism", col = "blue")

boxplot(df_nationalism$eth_vs_cult, main = "Boxplot of Ethnic vs Cultural Nationalism", ylab = "Ethnic - Cultural Nationalism", col = "lightgreen")


country_averages <- df_nationalism %>%
  group_by(countryname) %>%
  summarize(mean_eth_vs_cult = mean(eth_vs_cult, na.rm = TRUE),
            mean_l2immigrate = mean(l2immigrate, na.rm = TRUE),
            mean_l2hdi = mean(l2hdi, na.rm = TRUE),
            mean_nationalidentity = mean(nationalidentity, na.rm = TRUE),
            mean_l2gini = mean(l2gini, na.rm = TRUE))

# Histogram where each country equals 1
hist(country_averages$mean_eth_vs_cult,
     breaks = 10,  # You can adjust the number of bins
     main = "Histogram of Ethnic vs Cultural Nationalism by Country",
     xlab = "Mean Ethnic - Cultural Nationalism",
     col = "lightblue",
     border = "black")

```

###2. b) 2.- National identity

```{r}
summary(df_nationalism$nationalidentity)

sd(df_nationalism$nationalidentity)

hist(df_nationalism$nationalidentity, breaks = 20, main = "Histogram of National Identity", xlab = "National Identity Score", col = "lightblue")

# Create a weighted, stacked histogram with ggplot2
ggplot(df_nationalism, aes(x = nationalidentity, fill = countryname, weight = polweight)) +
  geom_histogram(bins = 20, color = "black") +
  scale_fill_manual(values = rainbow(length(unique(df_nationalism$countryname)))) +
  labs(title = "Weighted Histogram of National Identity Score",
       x = "Ethnic - Cultural Nationalism",
       y = "Weighted Frequency") +
  theme_minimal()


summary(df_nationalism$polweight)
plot(density(df_nationalism$nationalidentity), main = "Density Plot of National Identity", xlab = "National Identity Score", col = "blue")

boxplot(df_nationalism$nationalidentity, main = "Boxplot of National Identity", ylab = "National Identity Score", col = "lightgreen")

# Histogram where each country equals 1
hist(country_averages$mean_nationalidentity,
     breaks = 10,  # You can adjust the number of bins
     main = "Histogram of Mean National Identity by Country",
     xlab = "Mean National Identity",
     col = "lightblue",
     border = "black")

```

###2. b) 3.- Rejection of multiculturalism

```{r}
summary(df_nationalism$ougrcusttrad)

sd(df_nationalism$ougrcusttrad)

hist(df_nationalism$ougrcusttrad, breaks = 20, main = "Histogram of National Identity", xlab = "National Identity Score", col = "lightblue")

# Create a weighted, stacked histogram with ggplot2
ggplot(df_nationalism, aes(x = ougrcusttrad, fill = countryname, weight = polweight)) +
  geom_histogram(bins = 20, color = "black") +
  scale_fill_manual(values = rainbow(length(unique(df_nationalism$countryname)))) +
  labs(title = "Weighted Histogram of National Identity Score",
       x = "Ethnic - Cultural Nationalism",
       y = "Weighted Frequency") +
  theme_minimal()


summary(df_nationalism$polweight)
plot(density(df_nationalism$ougrcusttrad), main = "Density Plot of National Identity", xlab = "National Identity Score", col = "blue")

boxplot(df_nationalism$ougrcusttrad, main = "Boxplot of National Identity", ylab = "National Identity Score", col = "lightgreen")

# Histogram where each country equals 1
hist(country_averages$mean_ougrcusttrad,
     breaks = 10,  # You can adjust the number of bins
     main = "Histogram of Mean National Identity by Country",
     xlab = "Mean National Identity",
     col = "lightblue",
     border = "black")

```

### 2.c).- Regression plots: 
Using country averages to build level 2 regression plots showing the effect of 
the immigration rate on the mean difference between ethnic and cultural national 
identity scores and on the overall mean national identity score
```{r}
library(ggrepel)

# Create the scatter plot with country averages
ggplot(country_averages, aes(x = mean_l2immigrate, y = mean_eth_vs_cult)) +
  geom_point(color = "blue") +
  geom_smooth(method = "lm", formula = y ~ x, color = "darkgreen", se = FALSE) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2), color = "red", se = FALSE) +  # Add the regression line
  geom_text_repel(aes(label = countryname), 
                  box.padding = 0.35, 
                  point.padding = 0.5, 
                  segment.color = 'grey50') +
  # Tags for the regression lines
  annotate("text", x = 0, y = -1.3, 
           label = "y ~ x*ß", color = "darkgreen", hjust = 0, vjust = 1, fontface = "italic") +
  annotate("text", x = 0, y = -1.15, 
           label = "y ~ (x^2)*ß", color = "red", hjust = 0, vjust = 0, fontface = "italic") +
  labs(title = "Ethnic vs Cultural National Identity and Immigration Rate, by Country",
       x = "Immigration Rate",
       y = "Mean Ethnic - Cultural National Identity") #+
  #theme_minimal()


# Create the scatter plot with country averages
ggplot(country_averages, aes(x = mean_l2immigrate, y = mean_nationalidentity)) +
  geom_point(color = "blue") +
  geom_smooth(method = "lm", formula = y ~ x, color = "darkgreen", se = FALSE) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2), color = "red", se = FALSE) +  # Add the regression line
  geom_text_repel(aes(label = countryname), 
                  box.padding = 0.35, 
                  point.padding = 0.5, 
                  segment.color = 'grey50') +
  # Tags for the regression lines
  annotate("text", x = 25, y = 2.5, 
           label = "y ~ x*ß", color = "darkgreen", hjust = 0, vjust = 1, fontface = "italic") +
  annotate("text", x = 25, y = 2.3, 
           label = "y ~ (x^2)*ß", color = "red", hjust = 0, vjust = 0, fontface = "italic") +
  labs(title = "Overall National Identity and Immigration Rate, by Country",
       x = "Immigration rate",
       y = "Overall National Identity") #+
  #theme_minimal()

```

#3. Regression analysis
##3. a) Regression on difference between ethnic and cultural

```{r}
model_1 <- lm(eth_vs_cult ~ edulvl, data = df_nationalism, weights = polweight)

model_2 <- lm(eth_vs_cult ~ quintinc, data = df_nationalism, weights = polweight)

model_3 <- lm(eth_vs_cult ~ agerange, data = df_nationalism, weights = polweight)

model_4 <- lm(eth_vs_cult ~ edulvl + quintinc + agerange, data = df_nationalism, weights = polweight)

model_5 <- lm(eth_vs_cult ~ edulvl + quintinc + I(agerange^2), data = df_nationalism, weights = polweight)

stargazer(model_1, model_2, model_3, model_4, model_5,
          type = "text",  # "text" will print the table in the R console
          title = "Regression Results", 
          align = TRUE, 
          dep.var.labels = "Ethnic vs Cultural Nationalism",
          covariate.labels = c("Education Level", "Income Quintile", "Age Range"),
          no.space = TRUE)

```

##3. b) Regression on overall national identity

```{r}
model_1 <- lm(nationalidentity ~ edulvl, data = df_nationalism, weights = polweight)

model_2 <- lm(nationalidentity ~ quintinc, data = df_nationalism, weights = polweight)

model_3 <- lm(nationalidentity ~ agerange, data = df_nationalism, weights = polweight)

model_4 <- lm(nationalidentity ~ edulvl + quintinc + agerange, data = df_nationalism, weights = polweight)

model_5 <- lm(nationalidentity ~ edulvl + quintinc * agerange, data = df_nationalism, weights = polweight)

stargazer(model_1, model_2, model_3, model_4, model_5,
          type = "text",  # "text" will print the table in the R console
          title = "Regression Results", 
          align = TRUE, 
          dep.var.labels = "Overall National Identity Score",
          covariate.labels = c("Education Level", "Income Quintile", "Age Range"),
          no.space = TRUE)

```

#4. Mixed effects regression models
##4. a) Conduct mixed effects models (fixed slopes) on mediators:
###4. a) 1.- Level 2 effects on mediators (diff between ethnic and cultural AND overall national identity score)

```{r include=FALSE}
# Level 2 effects on difference between ethnic and cultural

print(" ")
print("Model 1: Linear Effect")
print(" ")

model_1 <- lmer(eth_vs_cult ~ l2immigrate + (1 | countryname), data = df_nationalism)

summary(model_1)

print(" ")
print("Model 2: Cuadratic effect")
print(" ")

model_2 <- lmer(eth_vs_cult ~ l2immigrate + I(l2immigrate^2) + (1 | countryname), data = df_nationalism)

summary(model_2)

# Level 2 effects on overall national identity level

print(" ")
print("Model 3: Linear Effect")
print(" ")

model_3 <- lmer(nationalidentity ~ l2immigrate + (1 | countryname), data = df_nationalism)

summary(model_3)

print(" ")
print("Model 4: Cuadratic effect")
print(" ")

model_4 <- lmer(nationalidentity ~ l2immigrate + I(l2immigrate^2) + (1 | countryname), data = df_nationalism)

summary(model_4)

# Create a tabular summary of the models
tab_model(model_1, model_2, model_3, model_4,
          title = "Level 2 Effects of Immigration Rate and HDI on Ethnic vs Cultural Nationalism",
          dv.labels = c("Model 1", "Model 2", "Model 3", "Model 4"),
          show.re.var = TRUE, # Show random effects variance
          show.icc = TRUE,    # Show Intraclass Correlation Coefficient
          show.aic = TRUE,    # Show AIC value
          show.r2 = TRUE,     # Show R-squared
          digits = 4,
          file = "C:/Users/oscar/Documents/Projects/Paper- Lv2EffectsOnNationalIdentity/Quantitative Analysis/Model fits/Crossnational - Regressions/ethvscult_regression_table.html")

```

##4. b) 2.- Conduct mixed effects models (fixed slopes) on outcome variable (rejecting multiculturalism):
Interacting eth_vs_cult (difference between ethnic and cultural national identity 
scores) with nationalidentity (overall national identity score across ethnic and 
cultural types), we can observe the differential effects of national identity and 
of whether this national identity is predominantly cultural or ethnic on negative 
attitudes towards immigrants

```{r include=FALSE}
print(" ")
print("Model 1")
print(" ")

model_1 <- lmer(ougrcusttrad ~ l2immigrate + (1 | countryname), data = df_nationalism)

summary(model_1)

print(" ")
print("Model 2")
print(" ")

model_2 <- lmer(ougrcusttrad ~ eth_vs_cult + nationalidentity + (1 | countryname), data = df_nationalism)

summary(model_2)

print(" ")
print("Model 3")
print(" ")

model_3 <- lmer(ougrcusttrad ~ eth_vs_cult * nationalidentity + (1 | countryname), data = df_nationalism)

summary(model_3)

print(" ")
print("Model 4")
print(" ")

model_4 <- lmer(ougrcusttrad ~ l2immigrate + eth_vs_cult * nationalidentity + (1 | countryname), data = df_nationalism)

summary(model_4)

# Create a tabular summary of the models
tab_model(model_1, model_2, model_3, model_4,
          title = "Level 2 and 1 effects on Rejecting Multiculturalism",
          dv.labels = c("Model 1", "Model 2", "Model 3", "Model 4"),
          show.re.var = TRUE, # Show random effects variance
          show.icc = TRUE,    # Show Intraclass Correlation Coefficient
          show.aic = TRUE,    # Show AIC value
          show.r2 = TRUE,     # Show R-squared
          digits = 4,
          file = "C:/Users/oscar/Documents/Projects/Paper- Lv2EffectsOnNationalIdentity/Quantitative Analysis/Model fits/Crossnational - Regressions/ougrcusttrad_regression_table.html")

```